Multiset-trie data structure

### 2022-07-29 tex/tex/s_data_structure.tex
Riste remarks (from the bus)

Let us have an example of a multiset-trie data structure. Let $\sigma = 2$ and 
$\Sigma = I = \{ 1,2 \}$ respectively, so the mapping $\phi$ is an identity mapping. 
Fix the degree of a node $n=3,$ so the maximal multiplicity of an element in 
a multiset is $n-1=2.$ The Figure~\ref{fig:sketch} presents the multiset-trie that 
contains multisets $\emptyset, \{ 1,1,2 \},$ $\{ 1,2,2 \},$ $\{ 2 \},$ $\{ 1,2 \},$ $\{ 2,2 \}.$ 
The \emph{Null} children are omitted on the figure.

\begin{figure}[h!]
\centering
\begin{tikzpicture}[>=stealth',
level 1/.style={sibling distance = 2.5cm},
level 2/.style={sibling distance = 1cm},
level distance = 1.5cm]
\node (R) [node1] {\tiny Root}
	child { node [node1] {} 
		child { node [node1] {} edge from parent node [near end, left] {$c_0$}}
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		child { node [node1] {} edge from parent node [near end, right] {$c_2$}}
		edge from parent node [near end, left] {$c_0$}	
	}
	child { node [node1] {}
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		child { node [node1] {} edge from parent node [near end, right] {$c_2$}}
		edge from parent node [near end, left] {$c_1$}
	}
	child { node [node1] {} 
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		edge from parent node [near end, right] {$c_2$} 
};	
%----------------- Level labels --------------	
	\begin{scope}[every node/.style={right}]
		\path (R 				-| R-3-1) ++(5mm,0) node {$L_1$};
		\path (R-3 			-| R-3-1) ++(5mm,0) node {$L_2$};
		\path (R-3-1		-| R-3-1) ++(5mm,0) node {$LL$};
	\end{scope}			
\end{tikzpicture}
\caption{Example of multiset-trie structure.}
\label{fig:sketch}
\end{figure}

>>>>

An example of a multiset-trie data structure with $\sigma = 2$ and 
$\Sigma = I = \{ 1,2 \}$ (i.e. the mapping $\phi$ is an identity mapping) is shown on Figure~\ref{fig:sketch}. 
On the figure, which stores elements of $\{\emptyset, \{ 1,1,2 \},$ $\{ 1,2,2 \},$ $\{ 2 \},$ $\{ 1,2 \},$ $\{ 2,2 \}\}$, the degree of a node is set to be $n=3,$ so the maximal multiplicity of an element in 
a multiset is $n-1=2.$

\begin{figure}[h!]
\centering
\begin{tikzpicture}[>=stealth',
level 1/.style={sibling distance = 2.5cm},
level 2/.style={sibling distance = 1cm},
level distance = 1.5cm]
\node (R) [node1] {\tiny Root}
	child { node [node1] {} 
		child { node [node1] {} edge from parent node [near end, left] {$c_0$}}
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		child { node [node1] {} edge from parent node [near end, right] {$c_2$}}
		edge from parent node [near end, left] {$c_0$}	
	}
	child { node [node1] {}
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		child { node [node1] {} edge from parent node [near end, right] {$c_2$}}
		edge from parent node [near end, left] {$c_1$}
	}
	child { node [node1] {} 
		child { node [node1] {} edge from parent node [near end, left] {$c_1$}}
		edge from parent node [near end, right] {$c_2$} 
};	
%----------------- Level labels --------------	
	\begin{scope}[every node/.style={right}]
		\path (R 				-| R-3-1) ++(5mm,0) node {$L_1$};
		\path (R-3 			-| R-3-1) ++(5mm,0) node {$L_2$};
		\path (R-3-1		-| R-3-1) ++(5mm,0) node {$LL$};
	\end{scope}			
\end{tikzpicture}
\caption{Example of multiset-trie structure containing multisets multisets $\emptyset, \{ 1,1,2 \},$ $\{ 1,2,2 \},$ $\{ 2 \},$ $\{ 1,2 \},$ $\{ 2,2 \}$. The \emph{Null} children are omitted.}
\label{fig:sketch}
\end{figure}

### 2022-07-29 tex/tex/s_math-noqed.tex
before plos submission
Riste remarks (from the bus)

\subsubsection{Mathematical model} \label{ss:mathmodel}
We start with the basics of our mathematical model. Let $\Sigma$ be an alphabet of
cardinality $\sigma,$ such that $\Sigma = \{ 1,2, \ldots, \sigma \}.$ Define $N$
to be the set of all possible multisets that can be inserted in
multiset-trie. Let $n$ be the maximal degree of a node in multiset-trie.
Then the maximal multiplicity of an element in a multiset is equal to $n-1.$
Thus, the number of multisets in a complete multiset-trie is $ |N| = n^{\sigma}.$
%
Let $M$ be a collection of multisets inserted into multiset-trie $\mathcal{M}.$
All the multisets in $M$ are constructed from the alphabet $\Sigma$ according
to the parameters $\sigma$ and $n$. Hence, any multiset $m\in M,$ 
has at most $\sigma$ distinct elements that are members of $\Sigma$ and 
every distinct element in $m$ has multiplicity strictly less than $n.$
%
Because a multiset does not distinguish different orderings, it is assumed, for
simplicity that all elements are ordered in an ascending order. A multiset $m$ is
represented as $\{1^{k_1},2^{k_2},\ldots, \sigma^{k_\sigma}\},$ where $e^{k_e}$
represents an element $e\in\Sigma$ with multiplicity $k_e.$
%

@@

\begin{lemma}\label{l:prob-node-existence}
Any potential node on a fixed level $i,$ where $i\in \{ 1,2,\ldots, \sigma +1 \}$ exists, %independently of others,
with probability
\begin{equation}
p_i=1-(1-p)^{n^{\sigma + 1 -i}}.
\end{equation}
\end{lemma}
\begin{proof}
Let $v$ be an arbitrary node in a multiset-trie on an arbitrary level $i.$ Consider
the subtree with the root $v$ and call it $v$-subtree. Since the height of the
multiset-trie is $\sigma + 1$ we can calculate the height of the $v$-subtree.
Taking in account that the root node has height 1, the height of the $v$-subtree is
%
\[
h_v = \sigma + 1 - i.
\]
%
A node in a multiset-trie exists if at least one node exists on the leaf level of its
subtree, i.e. a node on the level $\sigma + 1$ that belongs to $v$-subtree. The possible
number of nodes on the leaf level of $v$-subtree can be easily calculated knowing its height.
It is equal to
%
\[
n^{\sigma + 1 - i}.
\]
%
A node at level $\sigma+1$ exists with probability $p,$ where $p = P(m\in M).$
Thus, the probability that there are no nodes on leaf level in $v$-subtree is
%
\[
(1-p)^{n^{\sigma +1 - i}}.
\]
%
The claim follows by taking the complement probability of the above result. 
\hspace*{\fill}\qed
%
\end{proof}

@@

\begin{lemma}\label{l:prob-children}
Suppose that a node $v$ exists at level $1\leq i\leq\sigma$.
Then the number of its children $\xi_{iv}$ is modeled by a zero-truncated binomially
distributed random variable on parameters $n$ and $p_{i+1}$. In particular,
the probability of node $v$ having $k$ children equals to
\begin{equation}\label{eq:pdf}
P(\xi_{iv} = k) = \frac{\binom{n}{k} (1-p_{i+1})^{n-k}}{1-(1-p_{i+1})^n}
\end{equation}
and the corresponding probability generating function equals to
\begin{equation}\label{eq:generating_func}
G_i(z) = \frac{(1+p_{i+1}(z-1))^n - (1-p_{i+1})^n}{1-(1-p_{i+1})^n}.
\end{equation}
\end{lemma}
\begin{proof}
In order to prove the lemma, we have to show that
$\xi_{iv}\sim\mathcal{B}_0(n, p_{i+1}).$
Consider an arbitrary node $v$ on level $1\leq i\leq\sigma.$ According to the
definition of the multiset-trie a node exists at level $i$ if and only if
it has at least one child. Note that this is not true for the nodes on the leaf
level $\sigma + 1.$ Implies, a node on level $i$ can have
$k\in\{ 1,2,\ldots, n\}$ children. Let $X_0, X_1, \ldots, X_{n-1}$ be random
variables, they are defined as follows:
\[
X_k = \begin{cases}
0 & \textrm{child $k$ of node $v$ does not exist} \\
1 & \textrm{child $k$ of node $v$ exists} \\
\end{cases}
\]
As it was shown in previous Lemma~\ref{l:prob-children}, the distribution of 
$X_k$ is $X_k\sim Bernoulli(p_{i+1}).$ Since our model assumes that all the 
multisets in $M$ are chosen uniformly at random, the variables 
$X_k,X_l$ are independent for $k\neq l.$ But in our case the node $v$ can not 
have 0 children, so the sum $\sum_{k=1}^n X_k$ has a zero-truncated binomial 
distribution:
%
\[
\sum_{k=1}^n X_k \sim\mathcal{B}_0(n,p_{i+1})
\]
%
which completes the proof.
\hspace*{\fill}\qed
\end{proof}
%
Knowing the probability density and probability generating functions of $\xi_{ij}$ 
from Lemma~\ref{l:prob-children}, we now can estimate the number of nodes in 
a randomly generated multiset-trie as follows:
%
\begin{equation}\label{eq:num_nodes}
\mathbb{E}( | \mathcal{M} | ) = \mathbb{E}\left[ \sum_{i=1}^{\sigma+1} \xi_i \right].
\end{equation}
%

@@

\begin{proposition}\label{prop:exp_rec}
The expectation of the random variable $\xi_{i+1}$ can be expressed 
as follows.
\[
\mathbb{E}(\xi_{i+1}) = \mathbb{E}(\xi_i)\mathbb{E}(\mathcal{B}_0(n,p_{i+1}))
\]
for $1\leq i\leq\sigma.$
\end{proposition}
\begin{proof}
Using the following property of probability generating function
\begin{equation}\label{eq:prop_exp-pgf}
G_X'(1^-) = \mathbb{E}(X)
\end{equation}
the expectation for the random variable $\xi_{i+1}$ can be derived in terms of 
the equation~(\cref{eq:g_rec}).
\begin{eqnarray}\label{eq:exp}
\mathbb{E}(\xi_{i+1}) &=& G_{\xi_{i+1}}'(1^-) \nonumber \\
&=& G_{\xi_i}'(G_i(1^-))G_i'(1^-).
\end{eqnarray}
According to~(\cref{eq:pdf}) and~(\cref{eq:generating_func}) the value of $G_i(z)$ at 
$1$ is $1$ and the value of its derivative at $1$ is $\mathbb{E}(\mathcal{B}_0(n,p_{i+1})).$ 
Substituting the values of $G_i(1^-)$ and $G_i'(1^-),$ and applying the 
property~(\cref{eq:prop_exp-pgf}) we complete the proof.
\hspace*{\fill}\qed
\end{proof}
%
From the Proposition~\ref{prop:exp_rec} above and Lemma~\ref{l:prob-children} we 
can conclude that 
\begin{eqnarray}
\mathbb{E}(\xi_{i}) &=& \mathbb{E}(\xi_{i-1})\mathbb{E}\left( \mathcal{B}_0(n,p_{i}) \right) \nonumber \\
& = & \mathbb{E}(\xi_{i-1})\frac{np_{i}}{1-(1-p_{i})^n}.
\end{eqnarray}

\begin{theorem}\label{thm:exp_level}
Let $\mathcal{M}$ be a multiset-trie defined with parameters $n,$ $\sigma,$ and denote the number 
of nodes on every level $i$ by a random variable $\xi_i.$ Furthermore, let all multisets appear in 
$\mathcal{M}$ with equal probability $p\in (0,1).$ Then the expected number 
of nodes on every level of $\mathcal{M},$ i.e. $\mathbb{E}(\xi_i)$ is defined as 
\begin{equation}\label{eq:nodes_level}
\mathbb{E}(\xi_{i}) = n^{i-1} \frac{1-(1-p)^{n^{\sigma +1 -i}}}{1-(1-p)^{n^{\sigma}}}.
\end{equation}
\end{theorem}
\begin{proof}
According to~(\cref{eq:g_xi_one}) the expected number of nodes on the first level is 1. \\
Using $\mathbb{E}(\xi_1) = 1$ and the result from Proposition~\ref{prop:exp_rec}
we get
\begin{eqnarray*}
\mathbb{E}(\xi_{i}) &=& \prod_{j=2}^{i} \frac{n p_j}{1-(1-p_j)^n}
= \prod_{j=2}^{i} n \frac{1-(1-p)^{n^{\sigma +1-j}}}{1-(1-p)^{n^{\sigma + 2 -j}}} \\
&=& n^{i-1} \frac{1-(1-p)^{n^{\sigma +1 -i}}}{1-(1-p)^{n^{\sigma}}}
\end{eqnarray*}
\hspace*{\fill}\qed
\end{proof}

@@

\begin{theorem}\label{thm:exp_nodes}
The expected cardinality of a multiset-trie defined on parameters $n,$ $\sigma$ 
and $p$ can be computed as 
\begin{equation}
\mathbb{E}( | \mathcal{M} | ) = \sum_{i=1}^{\sigma+1} n^{i-1} \frac{1-(1-p)^{n^{\sigma +1 -i}}}{1-(1-p)^{n^{\sigma}}},
\end{equation}
where $r = (1-p)^n,$ so $r\in(0,1).$
\end{theorem}
\begin{proof}
Using the results obtained from Theorem~\ref{thm:exp_level} we compute 
\begin{eqnarray*}
\mathbb{E}( | \mathcal{M} | ) &=& \mathbb{E}\left[ \sum_{i=1}^{\sigma+1} \xi_i \right] \\
&=& \sum_{i=1}^{\sigma + 1} n^{i-1} \frac{1-(1-p)^{n^{\sigma + 1 -i}}}{1-(1-p)^{n^{\sigma}}} \\
\end{eqnarray*}
\hspace*{\fill}\qed
\end{proof}

@@

\begin{corollary}\label{cor:exp_nodes_param}
Let $M_1, M_2, \alpha_i$ and $\beta_i$ be defined as in previous Definition~\ref{def:params}, 
then the expected cardinality of a multiset-trie subtree $\mathcal{M}_{M_1}$ that contains all the multisets 
from the set $M_1$ is equal to 
\begin{equation}\label{eq:nodes_sub_param}
\mathbb{E}( |\mathcal{M}_{M_1}| ) = \sum_{i=1}^{\sigma + 1} \alpha_{i-1} \frac{1-(1-p)^{\alpha_{i-1}}}{1-(1-p)^{\alpha_{\sigma}}}.
\end{equation}
The expected cardinality of a multiset-trie subtree $\mathcal{M}_{M_2}$ that contains all the multisets 
from the set $M_2$ is equal to 
\begin{equation}\label{eq:nodes_super_param}
\mathbb{E}( |\mathcal{M}_{M_2}| ) = \sum_{i=1}^{\sigma + 1} \beta_{i-1} \frac{1-(1-p)^{\beta_{i-1}}}{1-(1-p)^{\beta_{\sigma}}}.
\end{equation}
\end{corollary}
%
\begin{proof}
Using the results from Theorem~\ref{thm:exp_level} and Theorem~\ref{thm:exp_nodes} 
we derive the formulas~(\cref{eq:nodes_sub_param}) and~(\cref{eq:nodes_super_param}) 
by specifying the possible number of nodes on every level in the multiset-trie according 
to the multiset $m.$ Note that the formula~(\cref{eq:nodes_level}) assumes that on every 
level but the first one there are $n$ possible nodes. Given submultiset or supermultiset 
query and an input multiset $m$ the number of nodes that will be traversed on level $i$ is 
defined by the number $k_{i-1}+1$ or $n-k_{i-1}-1$ for $i\geq 2.$ On level $i=1$ 
there is only one root node in any multiset-trie $\mathcal{M},$ which always exists 
if $M\neq\emptyset$ and is traversed for any type of query (submultiset and 
supermultiset).
\hspace*{\fill}\qed
\end{proof}

>>>>

\subsubsection{Mathematical model} \label{ss:mathmodel}
We start with the basics of our mathematical model. Let $\Sigma$ be an alphabet of
cardinality $\sigma,$ such that $\Sigma = \{ 1,2, \ldots, \sigma \}.$ Define $N$
to be the set of all possible multisets that can be inserted in a
multiset-trie. Let $n$ be the maximal degree of a node in a multiset-trie.
Then the maximal multiplicity of an element in a multiset is equal to $n-1.$
Thus, the number of multisets in a complete multiset-trie is $ |N| = n^{\sigma}.$
%
Let $M$ be a collection of multisets inserted into multiset-trie $\mathcal{M}.$
All the multisets in $M$ are constructed from the alphabet $\Sigma$ according
to the parameters $\sigma$ and $n$. Hence, any multiset $m\in M,$ 
has at most $\sigma$ distinct elements that are members of $\Sigma$ and 
every distinct element in $m$ has multiplicity strictly less than $n.$
%
Because a multiset does not distinguish different orderings, it is assumed, for
simplicity that all elements are ordered in an ascending order. A multiset $m$ is
represented as $\{1^{k_1},2^{k_2},\ldots, \sigma^{k_\sigma}\},$ where $e^{k_e}$
represents an element $e\in\Sigma$ with multiplicity $k_e.$
%

@@

\begin{lemma}\label{l:prob-node-existence}
Any potential node on a fixed level $i,$ where $i\in \{ 1,2,\ldots, \sigma +1 \}$ exists, %independently of others,
with probability
\begin{equation}
p_i=1-(1-p)^{n^{\sigma + 1 -i}}.
\end{equation}
\end{lemma}
\begin{proof}
Let $v$ be an arbitrary node in a multiset-trie on an arbitrary level $i.$ Consider
the subtree with the root $v$ and call it $v$-subtree. Since the height of the
multiset-trie is $\sigma + 1$ we can calculate the height of the $v$-subtree.
Taking in account that the root node has height 1, the height of the $v$-subtree is
%
\[
h_v = \sigma + 1 - i.
\]
%
A node in a multiset-trie exists if at least one node exists on the leaf level of its
subtree, i.e. a node on the level $\sigma + 1$ that belongs to $v$-subtree. The possible
number of nodes on the leaf level of $v$-subtree can be easily calculated knowing its height.
It is equal to
%
\[
n^{\sigma + 1 - i}.
\]
%
A node at level $\sigma+1$ exists with probability $p,$ where $p = P(m\in M).$
Thus, the probability that there are no nodes on leaf level in $v$-subtree is
%
\[
(1-p)^{n^{\sigma +1 - i}}.
\]
%
The claim follows by taking the complement probability of the above result.

%
\end{proof}

@@

\begin{lemma}\label{l:prob-children}
Suppose that a node $v$ exists at level $1\leq i\leq\sigma$.
Then the number of its children $\xi_{iv}$ is modeled by a zero-truncated binomially
distributed random variable on parameters $n$ and $p_{i+1}$. In particular,
the probability of node $v$ having $k$ children equals to
\begin{equation}\label{eq:pdf}
P(\xi_{iv} = k) = \frac{\binom{n}{k} (1-p_{i+1})^{n-k}}{1-(1-p_{i+1})^n}
\end{equation}
and the corresponding probability generating function equals to
\begin{equation}\label{eq:generating_func}
G_i(z) = \frac{(1+p_{i+1}(z-1))^n - (1-p_{i+1})^n}{1-(1-p_{i+1})^n}.
\end{equation}
\end{lemma}
\begin{proof}
In order to prove the lemma, we have to show that
$\xi_{iv}\sim\mathcal{B}_0(n, p_{i+1}).$
Consider an arbitrary node $v$ on level $1\leq i\leq\sigma.$ According to the
definition of the multiset-trie a node exists at level $i$ if and only if
it has at least one child. Note that this is not true for the nodes on the leaf
level $\sigma + 1.$ Implies, a node on level $i$ can have
$k\in\{ 1,2,\ldots, n\}$ children. Let $X_0, X_1, \ldots, X_{n-1}$ be random
variables, they are defined as follows:
\[
X_k = \begin{cases}
0 & \textrm{child $k$ of node $v$ does not exist} \\
1 & \textrm{child $k$ of node $v$ exists} \\
\end{cases}
\]
As it was shown in previous Lemma~\ref{l:prob-children}, the distribution of 
$X_k$ is $X_k\sim Bernoulli(p_{i+1}).$ Since our model assumes that all the 
multisets in $M$ are chosen uniformly at random, the variables 
$X_k,X_l$ are independent for $k\neq l.$ But in our case the node $v$ can not 
have 0 children, so the sum $\sum_{k=1}^n X_k$ has a zero-truncated binomial 
distribution:
%
\[
\sum_{k=1}^n X_k \sim\mathcal{B}_0(n,p_{i+1})
\]
%
which completes the proof.

\end{proof}
%
Knowing the probability density and probability generating functions of $\xi_{ij}$ 
from Lemma~\ref{l:prob-children}, we now can estimate the number of nodes in 
a randomly generated multiset-trie as follows:
%
\begin{equation}\label{eq:num_nodes}
\mathbb{E}( | \mathcal{M} | ) = \mathbb{E}\left[ \sum_{i=1}^{\sigma+1} \xi_i \right].
\end{equation}
%

@@

\begin{proposition}\label{prop:exp_rec}
The expectation of the random variable $\xi_{i+1}$ can be expressed 
as follows.
\[
\mathbb{E}(\xi_{i+1}) = \mathbb{E}(\xi_i)\mathbb{E}(\mathcal{B}_0(n,p_{i+1}))
\]
for $1\leq i\leq\sigma.$
\end{proposition}
\begin{proof}
Using the following property of probability generating function
\begin{equation}\label{eq:prop_exp-pgf}
G_X'(1^-) = \mathbb{E}(X)
\end{equation}
the expectation for the random variable $\xi_{i+1}$ can be derived in terms of 
the equation~(\cref{eq:g_rec}).
\begin{eqnarray}\label{eq:exp}
\mathbb{E}(\xi_{i+1}) &=& G_{\xi_{i+1}}'(1^-) \nonumber \\
&=& G_{\xi_i}'(G_i(1^-))G_i'(1^-).
\end{eqnarray}
According to~(\cref{eq:pdf}) and~(\cref{eq:generating_func}) the value of $G_i(z)$ at 
$1$ is $1$ and the value of its derivative at $1$ is $\mathbb{E}(\mathcal{B}_0(n,p_{i+1})).$ 
Substituting the values of $G_i(1^-)$ and $G_i'(1^-),$ and applying the 
property~(\cref{eq:prop_exp-pgf}) we complete the proof.

\end{proof}
%
From the Proposition~\ref{prop:exp_rec} above and Lemma~\ref{l:prob-children} we 
can conclude that 
\begin{eqnarray}
\mathbb{E}(\xi_{i}) &=& \mathbb{E}(\xi_{i-1})\mathbb{E}\left( \mathcal{B}_0(n,p_{i}) \right) \nonumber \\
& = & \mathbb{E}(\xi_{i-1})\frac{np_{i}}{1-(1-p_{i})^n}.
\end{eqnarray}

\begin{theorem}\label{thm:exp_level}
Let $\mathcal{M}$ be a multiset-trie defined with parameters $n,$ $\sigma,$ and denote the number 
of nodes on every level $i$ by a random variable $\xi_i.$ Furthermore, let all multisets appear in 
$\mathcal{M}$ with equal probability $p\in (0,1).$ Then the expected number 
of nodes on every level of $\mathcal{M},$ i.e. $\mathbb{E}(\xi_i)$ is defined as 
\begin{equation}\label{eq:nodes_level}
\mathbb{E}(\xi_{i}) = n^{i-1} \frac{1-(1-p)^{n^{\sigma +1 -i}}}{1-(1-p)^{n^{\sigma}}}.
\end{equation}
\end{theorem}
\begin{proof}
According to~(\cref{eq:g_xi_one}) the expected number of nodes on the first level is 1. \\
Using $\mathbb{E}(\xi_1) = 1$ and the result from Proposition~\ref{prop:exp_rec}
we get
\begin{eqnarray*}
\mathbb{E}(\xi_{i}) &=& \prod_{j=2}^{i} \frac{n p_j}{1-(1-p_j)^n}
= \prod_{j=2}^{i} n \frac{1-(1-p)^{n^{\sigma +1-j}}}{1-(1-p)^{n^{\sigma + 2 -j}}} \\
&=& n^{i-1} \frac{1-(1-p)^{n^{\sigma +1 -i}}}{1-(1-p)^{n^{\sigma}}}
\end{eqnarray*}

\end{proof}

@@

\begin{theorem}\label{thm:exp_nodes}
The expected cardinality of a multiset-trie defined on parameters $n,$ $\sigma$ 
and $p$ can be computed as 
\begin{equation}
\mathbb{E}( | \mathcal{M} | ) = \sum_{i=1}^{\sigma+1} n^{i-1} \frac{1-(1-p)^{n^{\sigma +1 -i}}}{1-(1-p)^{n^{\sigma}}},
\end{equation}
where $r = (1-p)^n,$ so $r\in(0,1).$
\end{theorem}
\begin{proof}
Using the results obtained from Theorem~\ref{thm:exp_level} we compute 
\begin{eqnarray*}
\mathbb{E}( | \mathcal{M} | ) &=& \mathbb{E}\left[ \sum_{i=1}^{\sigma+1} \xi_i \right] \\
&=& \sum_{i=1}^{\sigma + 1} n^{i-1} \frac{1-(1-p)^{n^{\sigma + 1 -i}}}{1-(1-p)^{n^{\sigma}}} \\
\end{eqnarray*}

\end{proof}

@@

\begin{corollary}\label{cor:exp_nodes_param}
Let $M_1, M_2, \alpha_i$ and $\beta_i$ be defined as in previous Definition~\ref{def:params}, 
then the expected cardinality of a multiset-trie subtree $\mathcal{M}_{M_1}$ that contains all the multisets 
from the set $M_1$ is equal to 
\begin{equation}\label{eq:nodes_sub_param}
\mathbb{E}( |\mathcal{M}_{M_1}| ) = \sum_{i=1}^{\sigma + 1} \alpha_{i-1} \frac{1-(1-p)^{\alpha_{i-1}}}{1-(1-p)^{\alpha_{\sigma}}}.
\end{equation}
The expected cardinality of a multiset-trie subtree $\mathcal{M}_{M_2}$ that contains all the multisets 
from the set $M_2$ is equal to 
\begin{equation}\label{eq:nodes_super_param}
\mathbb{E}( |\mathcal{M}_{M_2}| ) = \sum_{i=1}^{\sigma + 1} \beta_{i-1} \frac{1-(1-p)^{\beta_{i-1}}}{1-(1-p)^{\beta_{\sigma}}}.
\end{equation}
\end{corollary}
%
\begin{proof}
Using the results from Theorem~\ref{thm:exp_level} and Theorem~\ref{thm:exp_nodes} 
we derive the formulas~(\cref{eq:nodes_sub_param}) and~(\cref{eq:nodes_super_param}) 
by specifying the possible number of nodes on every level in the multiset-trie according 
to the multiset $m.$ Note that the formula~(\cref{eq:nodes_level}) assumes that on every 
level but the first one there are $n$ possible nodes. Given submultiset or supermultiset 
query and an input multiset $m$ the number of nodes that will be traversed on level $i$ is 
defined by the number $k_{i-1}+1$ or $n-k_{i-1}-1$ for $i\geq 2.$ On level $i=1$ 
there is only one root node in any multiset-trie $\mathcal{M},$ which always exists 
if $M\neq\emptyset$ and is traversed for any type of query (submultiset and 
supermultiset).

\end{proof}
